---
title: "Machine Learning Project"
author: "Daniel Brodie"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

This R Markdown document contains the code to train a model to predict exercise types from a collection of data captured from a weight lifting sample.  The data is available from <http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har>.

## Data Import and Cleaning

The code below loads the data and removes variables that have either near-zero variance, variables with many NA values, and variables that are not related to the activity capture (including, for example, the individual's name and timestamps).
```{r}

library(caret)
library(dplyr)

training <- read.csv("pml-training.csv")

#remove Near Zero Variance
nzv <- nearZeroVar(training,names=TRUE)
mydata <- select(training,-one_of(nzv))

#remove variables with many NA values
nav <- colnames(mydata)[colSums(is.na(mydata))>19000]
mydata <- select(mydata,-one_of(nav))

#remove non-data variables
nondata <- c("X","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","num_window","user_name")
mydata <- select(mydata,-one_of(nondata))

```

## Split data into training and testing datasets
``` {r}
inTrain = createDataPartition(mydata$classe,p=.75, list=FALSE)
sub.train <- mydata[inTrain,]
sub.test <- mydata[-inTrain,]

```
## Fit a Random Forrest Model

Note, this is using k-fold cross validation.  Also note, for speed sake, the code used to train the model is included below, but the model is loaded from disk.
``` {r}
train_control<- trainControl(method="cv", number=10, verboseIter = FALSE)
#mod1 <- train(classe~.,data=sub.train,method="rf", trControl=train_control)
load('model.rda')
```

## Estimate Out of Sample error rate by running sub.test against model
``` {r}
pred.sub <- predict(mod1,sub.test)
c1 <- confusionMatrix(pred.sub,sub.test$classe)
c1$overall[1]
```
